---
title: "Using HCbeta"
vignette: >
  %\VignetteIndexEntry{Using HCbeta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(hcinfer.use_emoji = TRUE)
```

HCbeta is the default estimator in `hcinfer()`. This vignette shows how to use
it, inspect the stored quantities, and run small sensitivity checks with its
method-specific arguments.

## Run HCbeta

Fit an OLS model and request HCbeta explicitly. This is equivalent to the
default `hcinfer(fit)` call.

```{r}
library(hcinfer)

schools <- PublicSchools |>
  dplyr::mutate(
    income_scaled = income / 10000,
    income_scaled_sq = income_scaled^2
  )

fit <- lm(expenditure ~ income_scaled + income_scaled_sq, data = schools)
result <- hcinfer(fit, type = "hcbeta")

summary(result)
```

## Extract coefficient test results

Use `tests()` to extract the coefficient-level Wald results as a tibble.

```{r}
tests(result)
```

For example, extract the robust standard error and p-value for the quadratic
income term.

```{r}
dplyr::select(tests(result, parm = "income_scaled_sq"), term, std_error, p_value)
```

## Inspect HCbeta parameters

HCbeta stores its six user-facing controls and its estimated quantities in
`method_params`. The adjustable controls are `c1`, `c2`, `lower`, `upper`,
`a_max`, and `b_max`. The fixed shape floor $\varepsilon = 0.01$ is not stored
as a control and is not accepted through `...`. It is applied after shrinkage
and is distinct from `lower`, which truncates the leverage complements.

```{r}
result$method_params
```

The table below distinguishes the six controls from the estimated diagnostics.
The first three rows cover all six adjustable values. The remaining rows report
quantities computed from the fitted design.

| Entry | Meaning |
|-------|---------|
| `c1`, `c2` | Constants controlling the HCbeta exponent. |
| `lower`, `upper` | Truncation limits for leverage complements. |
| `a_max`, `b_max` | Upper caps for the adjusted Beta shape parameters (default 10000, valid range `[50, 25000]`). |
| `a_tilde`, `b_tilde` | Adjusted Beta shape parameters. |
| `zeta` | Shrinkage weight used in the Beta parameter adjustment. |

These values are diagnostics. In routine use, the defaults should usually be
left unchanged. The fixed floor is absent because it is part of the HCbeta
definition rather than a sensitivity control.

## Inspect leverage and weights

HCbeta, like the other estimators, stores leverage values and robust weights.
This table shows the observations with the largest leverages.

```{r}
diagnostics <- tibble::tibble(
  observation = result$observation,
  leverage = result$leverage,
  weight = result$weights,
  residual = result$residuals
) |>
  dplyr::mutate(
    state = schools$state[as.integer(observation)],
    .before = leverage
  )

diagnostics |>
  dplyr::arrange(dplyr::desc(leverage)) |>
  dplyr::slice_head(n = 5)
```

You can also sort by robust weight.

```{r}
diagnostics |>
  dplyr::arrange(dplyr::desc(weight)) |>
  dplyr::slice_head(n = 5)
```

The covariance object can be plotted directly to display adjustment factors
against leverages.

```{r, fig.width = 7, fig.height = 4, fig.alt = "Scatterplot of HCbeta adjustment factors against leverage values for the public-schools model."}
plot(vcov_hc(fit, type = "hcbeta"))
```

## Use the covariance-only interface

Use `vcov_hc()` when you only need the HCbeta covariance matrix and diagnostics.

```{r}
hcbeta_cov <- vcov_hc(fit, type = "hcbeta")
hcbeta_cov
vcov(hcbeta_cov)
```

The covariance object stores the same method parameters and diagnostics.

```{r}
hcbeta_cov$method_params
summary(hcbeta_cov)
```

## Run a sensitivity check

All six adjustable HCbeta controls can be passed through `...`, and the two
shape caps can be supplied independently. The fixed shape floor cannot be
changed. A sensitivity check compares the default result with a small set of
alternative settings.

Each comparison should vary only declared controls so that its interpretation
remains tied to HCbeta. Changing `lower` or `upper` affects $w_t$ but leaves the
fixed shape floor unchanged. Passing `epsilon` is rejected rather than defining
a new estimator variant.

```{r}
sensitivity_results <- list(
  default = hcinfer(fit, type = "hcbeta"),
  c1_5 = hcinfer(fit, type = "hcbeta", c1 = 5),
  tighter_truncation = hcinfer(fit, type = "hcbeta", lower = 0.02, upper = 0.98),
  capped_shapes = hcinfer(fit, type = "hcbeta", a_max = 50)
)

sensitivity <- purrr::imap(sensitivity_results, \(res, setting) {
  row <- tests(res, parm = "income_scaled_sq")
  ci <- confint(res, parm = "income_scaled_sq")

  tibble::tibble(
    setting = setting,
    std_error = row$std_error,
    p_value = row$p_value,
    conf_low = ci$conf_low,
    conf_high = ci$conf_high,
    max_weight = max(res$weights)
  )
})
sensitivity <- dplyr::bind_rows(sensitivity)
sensitivity
```

This table helps identify whether the reported inference is sensitive to small
changes in HCbeta tuning constants. The defaults remain the recommended starting
point.
The fixed shape floor is omitted because it cannot be tuned.

## Compare HCbeta with one classical estimator

For a quick comparison, put HCbeta next to a classical estimator such as HC3.

```{r}
hc3 <- hcinfer(fit, type = "hc3")

extract_income <- function(res, method) {
  row <- tests(res, parm = "income_scaled_sq")
  ci <- confint(res, parm = "income_scaled_sq")
  tibble::tibble(
    method = method,
    estimate = row$estimate,
    std_error = row$std_error,
    p_value = row$p_value,
    conf_low = ci$conf_low,
    conf_high = ci$conf_high
  )
}

compare_hc3 <- dplyr::bind_rows(
  extract_income(result, "hcbeta"),
  extract_income(hc3, "hc3")
)
compare_hc3
```

## Practical workflow

For routine analyses:

1. Fit an OLS model with `lm()`.
2. Run `hcinfer(fit)` or `hcinfer(fit, type = "hcbeta")`.
3. Use `summary(result)` for readable output.
4. Use `tests(result)`, `confint(result)`, `coef(result)`, and `vcov(result)` for extracted values.
5. Inspect `result$weights`, `result$leverage`, and `result$method_params` when diagnostics are needed.

Use `vignette("hcinfer-comparison")` when you want to compare HCbeta with other
estimators implemented in the package.
