---
title: "Introduction to hcinfer"
vignette: >
  %\VignetteIndexEntry{Introduction to hcinfer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
options(hcinfer.use_emoji = FALSE)
```

This vignette gives a compact overview of `hcinfer`. It uses only base R for
data preparation, then shows the main functions, extraction methods, and plots.

## Data and Model

The examples use the `PublicSchools` data included with the package. The model
is the quadratic income model used in the HCbeta application.

```{r}
library(hcinfer)

schools <- PublicSchools
schools$income_scaled <- schools$income / 10000
schools$income_scaled_sq <- schools$income_scaled^2

fit <- lm(expenditure ~ income_scaled + income_scaled_sq, data = schools)
fit
```

The call to `lm()` omits observations with missing model variables.

## Available Estimators

Use `hc_methods()` to list the estimators and their default arguments.

```{r}
hc_methods()
```

The available types are `"hc0"`, `"hc1"`, `"hc2"`, `"hc3"`, `"hc4"`,
`"hc4m"`, `"hc5"`, `"hc5m"`, and `"hcbeta"`.

## Robust Inference

Use `hcinfer()` to compute robust inference for an `lm()` object. The default
estimator is HCbeta.

```{r}
result <- hcinfer(fit)
result
```

Use `summary()` for a readable report with coefficient tests, confidence
intervals, leverage diagnostics, and robust weight diagnostics.

```{r}
summary(result)
```

You can choose another estimator with `type`.

```{r}
result_hc3 <- hcinfer(fit, type = "hc3")
result_hc3
```

## Test Extraction

Use `tests()` to extract coefficient-level Wald tests as a tibble.

```{r}
tests(result)
```

Select one coefficient by name or position.

```{r}
tests(result, parm = "income_scaled_sq")
tests(result, parm = 3)
```

Change `alpha` to update the rejection decision. The test statistic and p-value
are not recomputed.

```{r}
tests(result, alpha = 0.10)
```

## Confidence Interval Extraction

Use `confint()` to extract robust confidence intervals.

```{r}
confint(result)
```

You can select a coefficient and change the confidence level.

```{r}
confint(result, parm = "income_scaled_sq")
confint(result, parm = "income_scaled_sq", level = 0.90)
```

## Coefficients and Covariance Matrices

The `coef()` method returns the OLS estimates stored in the result.

```{r}
coef(result)
```

The `vcov()` method extracts the robust covariance matrix stored in the
`hcinfer` object.

```{r}
robust_vcov <- vcov(result)
robust_vcov
```

Robust standard errors are the square roots of the diagonal entries.

```{r}
sqrt(diag(robust_vcov))
```

## Covariance-Only Workflow

Use `vcov_hc()` when you only need the robust covariance matrix and diagnostics.

```{r}
cov_hcbeta <- vcov_hc(fit)
cov_hcbeta
```

The same `summary()` and `vcov()` generics work for covariance objects.

```{r}
summary(cov_hcbeta)
vcov(cov_hcbeta)
```

You can also choose another estimator and pass its method constants.

```{r}
cov_hc5 <- vcov_hc(fit, type = "hc5", k = 0.7)
cov_hc5
```

## Plots

Use `plot()` on an `hcinfer` object to display robust confidence intervals.

```{r, fig.alt = "Robust confidence intervals for the public-schools regression coefficients."}
plot(result)
```

Select one coefficient with `parm`.

```{r, fig.alt = "Robust confidence interval for the quadratic income coefficient."}
plot(result, parm = "income_scaled_sq")
```

Use `plot()` on a `vcov_hc()` object to show leverage values and HC adjustment
factors.

```{r, fig.alt = "HCbeta adjustment factors plotted against leverage values for the public-schools regression."}
plot(cov_hcbeta)
```

Set `label_top` to control how many observations with the largest adjustment
factors are labeled.

```{r, fig.alt = "HC3 adjustment factors plotted against leverage values with the two largest weights labeled."}
plot(vcov_hc(fit, type = "hc3"), label_top = 2)
```

## A Small Comparison

The following base R code compares HCbeta and HC3 for the quadratic income
coefficient.

```{r}
test_hcbeta <- tests(result, parm = "income_scaled_sq")
test_hc3 <- tests(result_hc3, parm = "income_scaled_sq")

comparison <- data.frame(
  estimator = c("HCbeta", "HC3"),
  estimate = c(test_hcbeta$estimate, test_hc3$estimate),
  robust_se = c(test_hcbeta$std_error, test_hc3$std_error),
  p_value = c(test_hcbeta$p_value, test_hc3$p_value)
)

comparison
```

## Typical Workflow

For routine use, the workflow is short.

```{r, eval = FALSE}
fit <- lm(y ~ x1 + x2, data = data)
result <- hcinfer(fit, type = "hcbeta")

summary(result)
tests(result)
confint(result)
plot(result)
```
